---
title: "Review model outputs for tropical mountains"
author: JR Ferrer-Paris
output: html_document
editor_options: 
  chunk_output_type: console
---

Load programming environment variables
```{r}
here::i_am("review-RF-output/rf-models-and-predictions.qmd")
here::here("sandbox")
```

Load libraries
```{r}
require(dplyr)
require(sf)
require(raster)
library(caret)
library(stringr)
library(DT)
```

Load point records used to fit the model(s)

```{r}
records_file <- sprintf("%s/EFG-records-GMBA.rds",input.dir)
EFG_records <- readRDS(file=records_file) %>% st_drop_geometry()
all_units <- EFG_records %>% pull(GMBA_group) %>% unique()
```

Load the GMBA mountain inventory

```{r}

GMBA <- read_sf(sprintf("%s/GMBA_inventory_valid.gpkg", input.dir)) %>%
  filter(MapUnit %in% "Basic") 

```

Read existing files:
```{r}
completed_models <- 
  list.files(output.dir, 
           pattern = "rf-model-current.rda", 
           recursive = TRUE)
```



```{r}
GMBA_list <- 
  GMBA %>% 
    st_drop_geometry() %>% 
    filter(Level_03 %in% all_units) %>% 
    distinct(Level_01, Level_02, Level_03) %>%
    arrange_all() %>% 
    mutate(file_path=sprintf("GMBA_V2_L03-%s/rf-model-current.rda", 
                           str_replace_all(Level_03, "[ \\(\\)*\\/]+", "-")))
```

```{r}
completed_list <- GMBA_list %>% 
  filter(file_path %in% completed_models) %>% 
  dplyr::select(1:3)

datatable(completed_list)
```


We can use these tables to search for specific documentation for each mountain system:

```{r}
GMBA %>% 
  st_drop_geometry() %>% 
  filter(Level_03 %in% completed_list$Level_03) %>% 
  dplyr::select(GMBA_V2_ID, Level_01, Level_02, Level_03, MapName, WikiDataUR, Elev_High) %>%
  arrange(Level_01, Level_02, Level_03) %>%
  datatable()
  
```

For example, filtering to the Hawaian Islands, we found Haleakala as the first entry, we visited the corresponding wikidata site https://www.wikidata.org/wiki/Q515719, followed to the english wikipedia article for some brief overview. Then we searched google scholar with the search term "vegetation of HaleakalÄ" and found two entries (among many other):

- Mueller-Dombois, Dieter & Kitayama, K.. (1994). An altitudinal transect analysis of the windward vegetation on Haleakala, a Hawaiian island mountain: (2) vegetation zonation. Phytocoenologia. 24. 10.1127/phyto/24/1994/111. 
- Maaike Y. Bader, Max Rietkerk & Arnold K. Bregt (2007) Vegetation Structure and Temperature Regimes of Tropical Alpine Treelines, Arctic, Antarctic, and Alpine Research, 39:3, 353-364, DOI: 10.1657/1523-0430(06-055)[BADER]2.0.CO;2

We should prioritise some areas, for example start with Africa and then Asia. Focus on the best range (max elevations near 3000 to 5000)

Searching for "vegetation of Uluguru Mountains" in google scholar we found:

Burgess, N., Doggart, N., & Lovett, J. (2002). The Uluguru Mountains of eastern Tanzania: The effect of forest loss on biodiversity. Oryx, 36(2), 140-152. doi:10.1017/S0030605302000212


```{r}
EFG_records %>% 
  filter(GMBA_group %in% "Eastern Arc Mountains") %>% 
  group_by(source,EFG) %>% 
  summarise(records=n(), .groups = "drop")
```

